{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}

### 2016-12-31 proposal/lipservice.tex
write overview of mutexes
write overview pintos

The Pintos architecture \cite{pintos}, outlined in Section~\ref{sec:overview-pintos}, provides an opportunity to test Landslide's pedagogical mettle beyond CMU's walls.
For the experiments in Chapter~\ref{chap:quicksand}, I already implemented rudimentary support to test Pintos kernels.
Because Pintos provides substantially more basecode than Pebbles, most if not all of Landslide's kernel instrumentation can be done automatically,
using the names of the core scheduler functions already provided.

@@

\section{Transactional Memory}

>>>>

The Pintos architecture \cite{pintos}, outlined in Section~\ref{sec:overview-pintos}, provides an opportunity to test Landslide's pedagogical mettle beyond CMU's walls.
However, the existing scheduler and concurrency primitives which Pintos provides somewhat limits the remaining concurrency-critical code left to write,
hence limiting the degree to which Landslide can test student submissions.
This is in contrast to the Pebbles assignments at CMU, in which students must implement all the functionality listed in Section~\ref{sec:overview-pintos}.
On the other hand, the upside of Pintos providing substantially more basecode than Pebbles is that most, if not all, of Landslide's kernel instrumentation can be done automatically,
using the names of the core scheduler functions already provided.
For the experiments in Chapter~\ref{chap:quicksand}, I already implemented rudimentary support to test Pintos kernels, but more remains to be done.

@@

\section{Transactional Memory}
\label{sec:tm}

### 2016-12-31 proposal/overview.tex
write overview pintos

\subsection{Pintos}
\label{sec:overview-pintos}
% TODO

>>>>

\subsection{Pintos}
\label{sec:overview-pintos}

The Pintos kernel architecture \cite{pintos} is used at several universities, including Berkeley, Stanford, and the University of Chicago.
The Pintos basecode implements a rudimentary kernel, consisting of a context switcher, round-robin scheduler, locking primitives, and program loader.
upon which students add more features in several projects.
Most relevant to this thesis, the basecode provides the following functions/libraries, among others:
\begin{itemize}
	\item Semaphores (the basic concurrency primitive, implemented using direct scheduler calls): {\tt sema\_up}, {\tt sema\_down}, {\tt sema\_try\_down};
	\item Locks (which simply wrap a semaphore initialized to 1), {\tt lock\_acquire}, {\tt lock\_}- {\tt release}, {\tt lock\_try\_acquire};
	\item Condition variables (also implemented using scheduler calls): {\tt cond\_wait}, {\tt cond\_}- {\tt signal}, {\tt cond\_broadcast}, with the same semantics as Pebbles P2 condvars;
	\item Basic round-robin scheduling facilities: {\tt thread\_block} (a kernel-level analogue to Pebbles's {\tt deschedule}), {\tt thread\_yield}
	\item Kernel thread lifecycle management, {\tt thread\_create} and {\tt thread\_exit}, including stack space memory management;
	\item Interrupt and fault handlers;
	\item A page allocator, {\tt palloc\_get\_page}, {\tt palloc\_get\_multiple}, {\tt palloc\_}- {\tt free\_page}, {\tt palloc\_free\_multiple}
\end{itemize}
Both Pebbles and Pintos basecodes offer a standard C library including {\tt malloc}, string-formatting, printing, etc.

Although there is some variety in supplemental assignments, all Pintos courses include three core projects building on the Pintos basecode:
\begin{itemize}
	\item {\em Threads}: Students must implement an ``alarm clock'' (analogous to Pebbles's {\tt sleep} system call),
		a priority scheduling algorithm, and a multi-level feedback queue scheduler.
		% TODO: Later. Talk about how concurrency testing can only test certain parts of this crap.
	\item {\em Userprog}: Provided with rudimentary virtual memory and ELF loader implementations, students must implement argument passing and several system calls associated with userspace programs, including {\tt exec}, {\tt exit}, {\tt wait}, and file descriptor management.
	\item {\em Filesys}: Provided with a simple ``flat'' filesystem implementation, students must extend it with a buffer cache, extensible files, and subdirectories.
\end{itemize}

Some schools further offer a virtual memory project, extending the provided VM with a frame table and supplemental page table and fault handler \cite{standford-cs140,uchicago-cs230}, or supplemental HTTP server and {\tt malloc} assignments \cite{berkeley-cs162}.
Being largely architectural/algorithmic projects rather than concurrency-oriented ones, I am not concerned with these assignments in this thesis.
The main concurrency challenges in Pintos projects arise from the threads and userprog assignments:
implementing a correct {\tt alarm} routine,
ensuring the priority scheduler remains safe in the presence of concurrent threads of the same priority,
and designing correct interactions between the {\tt wait} and {\tt exit} system calls.
